Symbolic linearization of equations of motion of constrained multibody systems

### 2014-08-16 response_to_reviewers_aug_2014.txt

Response to Editor.

Dear Editor(s),
We feel that in this revision, as in the previous revisions, we have made an
effort to address as many of Reviewer #1's comments as possible. We have valued
a number of the comments provided by Reviewer #1, agreeing that the suggested
changes help to clarify the motivation and methods in the paper.

Unfortunately, there are 2 comments from Reviewer #1 we do not feel are
reasonable to address, comments 7 and 12.

Comment 7 simply states that the reviewer finds our presentation frustrating,
without offering specific faults, mistakes, unclear sections, areas that could
use improvement, etc.  Additionally, some of the other comments (comment 3) are
the reason we have included such a thorough description of Kane's method. There
seems to be a breakdown where the reviewer is not understanding how we are
applying this method, and it is not clear on our end which part of Section 2
needs to be improved to address the reviewer's difficulties.

Comment 12 states that the reviewer found the parts of the manuscript that they
understood confusing, and then admits to being unable to follow the derivations
we have presented. While we cannot simplify the math in the derivations, one
can always attempt to provide more clarification, but only if the reviewer
states which areas they had trouble with or where throughout the derivation
they became lost. This comment as given to us, does not allow us to do
anything.

In summary, we have addressed as many of the comments as possible, but feel
that the other comments are not reasonable, and suggest that Reviewer #1 has
not made a good faith effort to either comprehend the paper, or communicate
where they had difficulty comprehending the paper. As Reviewer #2 seems to
approve of the manuscript, we would ask that you overlook Reviewer #1's
concerns when making the decision to publish or reject this work.

Thank you,
Luke Peterson
Gilbert Gede
Mont Hubbard

Response to Reviewer #1.

1. Abstract. “...but is compatible with any method as long as only ordinary
    differential equations are required to describe the system, and independent
    and dependent states remain in the final equations (i.e. no DAE's,
    dependent variables are not substituted away)”. The sentence, and
    especially its end, is still not clear to me.

Author Response: We have changed the sentence for clarity. It now more clearly
specifies the types of differential equations which can be used.  It also
clarifies that it was developed for use with Kane's Method when dependent
quantities have not been algebraically eliminated.

2. 1. 36-37 (page 1, lines 36-37). “Design of linear state feedback controllers
    and first order stability analysis of equilibria are two of the most common
    uses of multibody system dynamics models”. I am not sure about these are
    the most common uses.

Author Response: We removed the words which suggested that these were the most
common uses, instead stating that they are just common uses.

3. 2. 21-19. The paragraph is still unclear to me. Kane’s equations are the
   projections of the initial (expressed in dependent velocities and as such
   constraint reaction-induced) dynamic equations into the directions of
   generalized speeds, which are by assumption independent for a constrained
   system, and with these projections the constraint reactions are excluded
   from evidence. If the projected dynamic equations are still expressed in the
   dependent velocities (not in the generalized speeds), the number of the
   projected dynamic equations is smaller than the number of the velocities. In
   this context I do not understand the remark that “The kinematical and
   dynamical differential equations can be solved for the time derivatives of
   the generalized coordinates and generalized speeds, ...” as well as most of
   the content of the paragraph. My reservations raised in the previous review
   were not clarified.

Author Response: We believe you have some confusion regarding the different
possible ways to apply Kane's method, which is why we Section 2 clarifies how
we use Kane's method. One possible way to apply the method is when a
non-minimal set of speeds are used to define the velocities of the system; in
this case, the velocity constraint matrix B (equation 11 in the paper) can be
reworked into matrix A (equation 21). This A matrix is used to reduce the
number of equations for generalized speeds by the number of constrained speeds.
This is also outlined by Kane in [Kane 1985], in equations 4.4.3 and 4.4.11,
which shows the reduction in size of the equations for the generalized speeds
and that _all_ the speeds (dependent and independent) are still present.

4. 2. 20-27. I am not totally satisfied with the changes/clarifications made by
    the Authors. For example, the remark “Velocity constraints often appear
    where there is rolling without slip or there are closed kinematic loops” is
    unclear (if correct at all). Velocity constraints, or, better, constraint
    equations at the velocity level, are either nonholonomic constraints (often
    consequent to the rolling without slip) or differentiated with respect to
    time holonomic constraint equations (one source of such constraints are
    closed kinematic loops).

Author Response: We have rewritten this entire paragraph to ensure that the
description of each level of constraints is consistent, and more clear. We have
reworked the sentence in question to more clearly state that things like
rolling without slip lead to nonholonomic constraints and closed kinematic
loops lead to time differentiated holonomic constraints, both of which are
velocity level constraints.

5. 5. 25-50. The velocity constraints are introduced in Eq. (8) in a general
    (nonlinear) form f_v (q,u,t) = 0 . It is then said that “By assumption (why
    “by assumption”?), the velocity constraints are linear with respect to u,
    ... which implies Bu + fvt(q,t) = 0”. Firstly, in general, the constraint
    matrix B is not constant, i.e.  B(q) or even B(q,t). Then, why the velocity
    constraints were not introduced from the beginning in this “linear” form?

Author Response: We have clarified that the B matrix is a function of the
coordinates and time, even though it is written as only a single symbol (a
format which we use throughout the paper).

6. 8. 15 and 831. The meaning of the “exogenous inputs” is not well
    motivated/explained.

Author Response: We believe that the sentence "The s exogenous inputs include
control system inputs and/or external disturbance forces or torques applied to
the system." explains that the exogenous inputs include forces/torques which
are not internal to the system (e.g. a reaction force), but instead applied
from outside the system (a motor driving a body in the system).

7. A general remark to Section 2. While Kane’s method is generally well
    understood, the Authors presentation of its formulation is rather
    “frustrating”.

8. 10. 32-34. One of the velocity constraint equations introduced in Eq. (30)
    should be a time differentiated form of the configuration constraint of Eq.
    (25). It seems that it is Eq. (30c).  If so, it should rather read 
    r u2 s2 + u6  = 0 (not r u2 s3 + u6 = 0). Can you explain this?

Author Response: While it could be written this way, what we have presented is
correct. The method you are suggesting requires ^N v ^P* to be expressed in N.
Then, the N_x and N_y components would be 2 of the motion constraints, and the
time differentiated configuration constraint will equal the N_z component. This
can be simply verified by time differentiating the configuration constraint and
utilizing equation 34 to elimante the q_dots, so only u's are present. In
comparing this to the projection of ^N v ^P* in the N_z direction, you will
find these are equivalent. We believe what we have presented is simpler.

9. 10. 36-38. It is written: “These equations are linear in the ui terms,
    nonlinear in the qi terms, involve geometric system parameters (but not
    mass or inertial parameters), and do not explicitly involve \dot{qi}
    terms”. The remark that constraint equations do not involve inertial
    parameters is a cliché. According to Eq. (11), \dot{qi} terms are neither
    involved.

Author Response: Although it might be cliché, we feel explicitness is valuable,
and we do not wish to remove a sentence which might help a reader avoid
confusion.

10. 10. 39-118. The paragraph is unintelligible. The “diabolical set of
    coordinates and speeds” is just a set of dependent coordinates and speeds,
    resulted in one configuration constraint of Eq. (25) and three velocity
    constraints of Eq. (30) (one holonomic constraint of Eq.  (25) and two
    nonholonomic constraints of Eq. (30a) and (30b)). There are then five
    independent coordinates and three independent speeds. By choosing the
    independent coordinates and speeds, which may in general not be subsets of
    qi and ui, respectively, the constraint equations (25) and (30) are
    satisfied by identity.

Author Response: We have changed "diabolical" to "purposefully nonminimal".
That should clairfy that the example exists to show the technique on a
tractable system with a nonminimal set of coordinates and speeds, and to
highlight the approach, rather than suggest that we should be using these
coordinates and speeds on the rolling disk example.

11. 11. 14. Is Eq. (31c) correct? (see Remark 8).

Author Response: Yes it is correct, our response to comment 8 addresses this.
The code in the electronic supplementary material can be used to verify this.

12. A general remark. The complexity of the followed parts of the manuscript is
    not appealing to me. In fact, I was unable to follow the derivations. I may
    suppose that the same difficulties will have most of the potential readers
    of the paper, not mentioning those who would like to follow the procedures.

### 2014-08-29 PetersonGedeHubbard2012.tex
Implement Mont's feedback from 3rd round of revisions

\section{Introduction}
\label{sec:intro}
Design of linear state feedback controllers and first order stability analysis
are common uses of multibody system dynamics models. Both require linearized
equations of motion. For unconstrained systems, obtaining the linearized
equations of motion is as simple as arranging them  in first order form
($\dot{x} = f(x, t)$) and evaluating $\frac{\partial f}{\partial x}}$ at the
point of interest.  However, when the system has constraints not all the states
are independent, and this process is not as straightforward.

For a system with constraints, Kane's dynamical equations\cite{Kane1985} are
first order ODEs (in the generalized speeds) which include both the
independent and dependent states (assuming the dependent states have not been
algebraically eliminated). The kinematical and dynamical differential equations
can be solved for the time derivatives of the generalized coordinates and
generalized speeds, respectively. Taken together, it is tempting to think that
this system of first order differential equations could be linearized in the
same fashion as an unconstrained system.  Unfortunately, this is not the case.
Linearizing these equations in the same manner as one would for an
unconstrained system (by directly computing the Jacobian) is incorrect and will
yield incorrect linearized equations (and subsequently incorrect eigenvalues,
in the case of a stability analysis). We illustrate this by means of a familiar
example, the rolling disk, formulated with a non-traditional set of coordinates
and generalized speeds which are not all independent. We then provide a
systematic method to correctly generate the linear equations of motion for
constrained systems which have been found using Kane's method (or any other
method that results in equations with the same form). We apply this method to
the rolling disk example to illustrate that it yields eigenvalues that match
previously published results.

@@

% TODO: Improve literature review.
Bottema appears to be the first author to have shown that special
considerations are needed for linearizing nonholonomic systems
\cite{Bottema1949}. While other authors have examined linearization of
nonholonomic systems, we have found that most techniques are not directly
applicable to equations formed using Kane's method. While we do not doubt that
they \textit{could} be applied correctly, the details of how to do so, as far
as the authors are aware, have not been explicitly presented. Kang et al.
\cite{Kang2003} and Negrut and Ortiz \cite{Negrut2006} have both explored
linearization, but their methods have been developed for equations of motion
found using Lagrange's method, which contain Lagrange multipliers ( not present
in equations of motion generated using Kane's method). Further, both of these
papers lack a discussion of basic concepts such as how many quantities are
independent, and in the case of eigenvalue computation, how many eigenvalues
should be present. Minaker and Rieveley \cite{Minaker2010} and Schwab and
Meijaard \cite{Schwab2003} have developed techniques for generating equations
of motion (and linear forms thereof) using Lagrange's method; the work of
Schwab and Meijaard can be applied symbolically while Minaker and Reiveley's
approach is numeric in nature. Neither discusses the choice of generalized
speeds as in Kane's method: both form the equations of motion in terms of
second time derivatives of the coordinates.  Further, neither of these works
addresses constraints which cannot be eliminated (e.g., nonlinear configuration
constraints), nor do they address what approach should be taken when a single
choice of dependent states may not suffice.  Neuman \cite{Neuman1984} presents
a symbolic (as opposed to numeric) formulation for linearization of the
Q-matrix formulation of the Lagrangian, but makes no mention of constrained
systems and attempts to contact the author regarding the software (algebraic
robot modeler -- Arm) as well as internet searches for the software have been
fruitless.

@@

\section{Kane's Method, briefly}
\label{sec:kane_method}
To familiarize the reader with Kane's method, several concepts relating to its
use are discussed: the use of generalized speeds in addition to generalized
coordinates; how these generalized speeds can be used to project permissible
motions of the system, and how this removes the need to consider
non-contributing (internal constraint) forces.  Then, the mathematical steps
required to use Kane's method are illustrated.

In Lagrange's method, generalized coordinates describe the configuration of a
system while time derivatives of the generalized coordinates characterize the
velocities of a system.  Kane's method allows velocities to be written in terms
of generalized speeds, which are not required to be the time derivatives of
coordinates (although such a definition is permitted).

@@

Taking the dot product of each partial velocity with both sides of Newton's
second law, $\bar{F}=m\bar{a}$ (or the rotational equivalent), only terms that
are parallel to the partial velocities remain; one scalar equation is generated
for each generalized speed. An important byproduct is that internal constraint forces
(non-contributing forces) imposed by objects such as pin or sliding joints no
longer need to be considered. The reason for this is that partial velocities
are \textit{by construction} perpendicular to forces of constraint. Constraint
forces can be included in the formulation, but will not appear in the final
equations of motion because of this very desirable property.

In summary, Kane's method allows for velocities to be defined in terms of
generalized speeds, allowing for simpler velocity expressions.  The dot product
of each partial velocity (the direction of motion associated with each
generalized speed) with Newton's second law (or Euler's equations) removes all
terms which are not related to the rate of change of each generalized speed.
This generally results in simple equations of motion, which are easier to form.

@@

\section{Dynamics Example}
\label{sec:example}
To illustrate Kane's method in the presence of configuration and velocity
constraints, we consider a thin disk $C$ rolling without slip on a horizontal
plane. The dextral unit vectors $\hat{c}_x, \hat{c}_y, \hat{c}_z$ are fixed to $C$
with $\hat{c}_y$ perpendicular to the disk plane. Let $r$ and $m$ be the
disk radius and mass, and assume a thin disk mass distribution so that
$I_{xx}=I_{zz}=mr^2/4$ and $I_{yy} = mr^2/2$. Take the inertial (Newtonian)
frame $N$ to be similarly equipped with dextral unit vectors $\hat{n}_x,
\hat{n}_y, \hat{n}_z$, with $\hat{n}_z$ perpendicular to the ground plane
(downwards) and aligned with the local gravitational field. To orient the disk,
first align $C$ with $N$, then apply successive body fixed ZXY rotations:
yaw $q_1$, roll $q_2$, and spin $q_3$.  Denote by $A$ and $B$ the two frames
which persist after the first two intermediate rotations in the sequence; we
refer to these as the instantaneous yaw and lean frames, respectively. The unit
vector, $\hat{b}_z$ fixed in the $B$ frame, is parallel to the disk plane and
aligned with the position vector from the disk center $C^*$ to the lowest point
of the disk $P^*$. Figure \ref{fig:rolling_disc} shows how the generalized
coordinates are used to configure the disc.

@@

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.8\textwidth]{disk_drawing_cropped.pdf}
    \caption{Configuration of rolling disc using non-minimal coordinates.}
    \label{fig:rolling_disc}
\end{figure}

@@

At this point, the reader familiar with the rolling disk problem may wondering
why we have purposefully chosen a non-minimal set of coordinates and speeds.
In this simple example, there is an {\bf \textit{obvious and minimal}} choice
of coordinates and speeds ($q_6, u_4, u_5, u_6$, should be dependent; this
choice will be valid for all parameters and configurations; further, the
derivation of the equations of motion can be done without ever introducing
these quantities), for some systems this is not the case (e.g., Stewart's
platform, bicycle). The numeric values of system parameters and the
configurations in which the system will operate will determine which
coordinates and speeds can be taken to be dependent (and hence potentially
eliminated by algebraic manipulations).  However, it may not be possible
\textit{a priori} to determine easily which coordinates and speeds should be
taken as dependent; further, some systems may operate in distinct enough
regions of the configuration space that one \textit{must} use different choices
of dependent coordinates and/or speeds as the system moves from one region to
another. The purpose of our complex choice of coordinates and speeds is to
illustrate, in the context of a familiar example, how constraints must be taken
into account when linearizing equations of motion. It is our hope that by doing
this for a well known and relatively simple example, readers can apply the same
techniques to more complicated systems where its use is more appropriate or
necessary.

@@

The kinematic differential equations relate the time derivatives of the
coordinates and the generalized speeds. They are obtained by equating the
velocities written using time-differentiated generalized coordinates and the
velocities written using generalized speeds. From the addition theorem for
angular velocity, the angular velocity of $C$ is
\begin{equation}
  \label{w_qdot}
  {^N}\bar{\omega}^C = \dot{q}_1 \hat{n}_z + \dot{q}_2 \hat{a}_x + \dot{q}_3 \hat{b}_y
\end{equation}
while time differentiating $\bar{r}^{C^*}$ in $N$ yields
\begin{equation}
  \label{v_qdot}
  {^N}\bar{v}^{C^*} = \dot{q}_4 \hat{n}_x + \dot{q}_5 \hat{n}_y + \dot{q}_6 \hat{n}_z
\end{equation}
Equating (\ref{w_u}) with (\ref{w_qdot}) and (\ref{v_u}) with (\ref{v_qdot}),
and resolving these vector equations into the $B$ frame, and rearranging so
that all terms appear to the left of the equality sign, we obtain
\begin{align}
    \label{rd:f_0_f_1}
\underbrace{\left[\begin{matrix}\dot{q}_{2}\\s_2
    \dot{q}_{1} + \dot{q}_{3}\\ c_2
    \dot{q}_{1}\\\dot{q}_{4}\\\dot{q}_{5}\\\dot{q}_{6}\end{matrix}\right]}_{\mathbf{f}_0}
    +
\underbrace{\left[\begin{matrix}- u_{1}\\- u_{2}\\- u_{3}\\
\left(s_1 s_2 s_3 - c_1c_3\right)u_4 +
s_1 c_2 u_5 - \left(s_1 s_2c_3 +
s_3c_1\right)u_6 \\
%
\left(-s_1c_3 - s_2 s_3c_1\right)u_4 -
c_1c_2 u_5 +
\left(-s_1 s_3 + s_2c_1c_3\right)u_6
\\
%
s_3c_2u_4  - s_2u_5 - c_2c_3u_6
\end{matrix}\right]}_{\mathbf{f}_1}
    = \left[\begin{matrix} 0\\ 0\\ 0\\ 0\\ 0\\ 0\end{matrix}\right]
\end{align}
The translational acceleration of $C^*$ and the angular acceleration of $C$,
relative to $N$, are
\begin{align}
    {^N}\bar{a}^{C^*} =& (-(u_1 s_3 + u_3 c_3) u_5 + u_2 u_6 +
                         \dot{u}_4) \hat{c}_x \notag \\
                      & + ((u_1 s_3 + u_3 c_3) u_4 - (u_1
                         c_3 - u_3 s_3) u_6 + \dot{u}_5) \hat{c}_y
                         \notag\\
                      & + ((u_1 c_3 - u_3 s_3) u_5 - u_2 u_4 +
                         \dot{u}_6) \hat{c}_z \\
    {^N}\bar{\alpha}^{C} =& (-u_2 u_3 + u_3^2 t_2 + \dot{u}_1) \hat{b}_x
                         + \dot{u}_2 \hat{b}_y + (u_1 u_2 - u_1 u_3 t_2 +
                         \dot{u}_3) \hat{b}_z
\end{align}

@@

The generalized active forces for this example are
\begin{align}
    \bar{R}_{C^*} &= m g \hat{n}_z \\
    \bar{T}_C &= 0
\end{align}

@@

The twelve unknown quantities appearing in the nine kinematic and dynamic
differential equations are $\dot{q}_{1-6}$ and $\dot{u}_{1-6}$. The three
acceleration constraint equations provide the final three equations which allow
for all twelve unknowns to be solved. Combining (\ref{rd:f_0_f_1}),
(\ref{rd:f_2_f_3}), and (\ref{rd:f_a}), the complete set of motion equations
can be written.
\begin{align}
\label{rd:ode}
\begin{bmatrix} \mathbf{f}_0 + \mathbf{f}_1\\
                \mathbf{f}_2 + \mathbf{f}_3\\
                \mathbf{f}_a \end{bmatrix} = \mathbf{0}
\end{align}

The naive approach to linearizing these equations of motion would be to solve
for the ($\dot{q}_i$, $\dot{u}_i$) terms to construct the right-hand side of
\begin{align}
\begin{bmatrix}\dot{\mathbf{q}} \\ \dot{\mathbf{u}}\end{bmatrix} =
    \mathbf{f}(\mathbf{q}, \mathbf{u}, t)
\end{align}
and take the Jacobian of $\mathbf{f}$ with respect to $[\mathbf{q} \quad
\mathbf{u}]^T$. However, this will give incorrect results. As discussed in
\cite{Schwab2003}, when $m, r,$ and $g$ are taken to be unity, and the system
is linearized about the upright steady rolling condition, the critical speed is
$v \triangleq -r\dot{q}_3 =\pm\frac{1}{\sqrt{3}}$. When following the naive approach, and
evaluating the Jacobian at the same operating conditions and parameters, eight
of the twelve eigenvalues are identically zero while the remaining four are:
\begin{align}
  \label{eq:evals_incorrect}
  \lambda_{1,2}=\pm\frac{\sqrt{6}}{3}\sqrt{-v^2},
%  \quad \lambda_{3,4} = \pm\sqrt{\frac{4}{5} - \frac{12}{5} v^2}
  \quad \lambda_{3,4} &= \pm \frac{2\sqrt{1 - 3v^2}}{\sqrt{5}}
\end{align}

@@

Further, the fact that twelve eigenvalues are obtained should be an alert that
something is incorrect, since the number of independent quantities is only
eight (five coordinates and three speeds). We now outline a linearization
procedure which addresses this issue; we revisit this example in Section
\ref{sec:example_revisited}.

>>>>

\section{Introduction}
\label{sec:intro}
Design of linear state feedback controllers and first order stability analysis
are common uses of multibody system dynamics models. Both require linearized
equations of motion.  We define a constrained system to be one whose
configuration has been described by a non-minimal set of generalized
coordinates or whose motion has been described by a non-minimal set of
generalized speeds.  For unconstrained systems (with a minimal set of
generalized coordinates and speeds), obtaining the linearized equations of
motion is as simple as arranging them in first order form ($\dot{x} = f(x, t)$)
and evaluating $\frac{\partial f}{\partial x}}$ at the point of interest.
However, when the system has constraints not all the states (coordinates and
speeds) are independent, and this process is not as straightforward.

For a system with constraints, Kane's dynamical equations\cite{Kane1985} are
first order ODEs (in the generalized speeds) which include both the independent
and dependent states (assuming the dependent states have not been algebraically
eliminated). The kinematical differential equations can be solved for the time
derivatives of the generalized coordinates and the dynamical differential
equations can be solved for the time derivatives of the generalized speeds.
Taken together, it is tempting to think that this system of first order
differential equations could be linearized in the same fashion as an
unconstrained system.  Unfortunately, this is not the case.  Linearizing these
equations in the same manner as one would for an unconstrained system (by
directly computing the Jacobian) is incorrect and will yield incorrect
linearized equations (and subsequently incorrect eigenvalues, in the case of a
stability analysis). We illustrate this by means of a familiar example, the
rolling disk, formulated with a non-traditional set of coordinates and
generalized speeds which are not all independent. We then provide a systematic
method to correctly generate the linear equations of motion for constrained
systems which have been found using Kane's method (or any other method that
results in equations with the same form). We apply this method to the rolling
disk example to illustrate that it yields eigenvalues that match previously
published results.

@@

% TODO: Improve literature review.
Bottema appears to be the first author to have shown that special
considerations are needed for linearizing nonholonomic systems
\cite{Bottema1949}. While other authors have examined linearization of
nonholonomic systems, we have found that most techniques are not directly
applicable to equations formed using Kane's method. While we do not doubt that
they \textit{could} be applied correctly, the details of how to do so, as far
as the authors are aware, have not been explicitly presented. Kang et al.
\cite{Kang2003} and Negrut and Ortiz \cite{Negrut2006} have both explored
linearization, but their methods have been developed for equations of motion
found using Lagrange's method, which contain Lagrange multipliers (not present
in equations of motion generated using Kane's method). Further, both of these
papers lack a discussion of basic concepts such as how many quantities are
independent, and in the case of eigenvalue computation, how many eigenvalues
should be present. Minaker and Rieveley \cite{Minaker2010} and Schwab and
Meijaard \cite{Schwab2003} have developed techniques for generating equations
of motion (and linear forms thereof) using Lagrange's method; the work of
Schwab and Meijaard can be applied symbolically while Minaker and Reiveley's
approach is numeric in nature. Neither discusses the choice of generalized
speeds as in Kane's method: both form the equations of motion in terms of
second time derivatives of the coordinates.  Further, neither of these works
addresses constraints which cannot be eliminated (e.g., nonlinear configuration
constraints), nor do they address what approach should be taken when a single
choice of dependent states may not suffice.  Neuman \cite{Neuman1984} presents
a symbolic (as opposed to numeric) formulation for linearization of the
Q-matrix formulation of the Lagrangian, but makes no mention of constrained
systems and attempts to contact the author regarding the software (algebraic
robot modeler -- Arm) as well as internet searches for the software have been
fruitless.

@@

\section{Kane's Method, briefly}
\label{sec:kane_method}
To familiarize the reader with Kane's method, several concepts relating to its
use are discussed: the use of generalized speeds in addition to generalized
coordinates; how these generalized speeds can be used to project permissible
motions of the system; and how this removes the need to consider
non-contributing (internal constraint) forces.  Then, the mathematical steps
required to use Kane's method are illustrated.

In Lagrange's method, generalized coordinates describe the configuration of a
system while time derivatives of the generalized coordinates characterize the
velocities.  Kane's method allows velocities to be written in terms
of generalized speeds, which are not required to be the time derivatives of
coordinates (although such a definition is permitted).

@@

Taking the dot product of each partial velocity with both sides of Newton's
second law, $\bar{F}=m\bar{a}$ (or the rotational equivalent), only terms that
are parallel to the partial velocities remain; one scalar equation is generated
for each generalized speed. An important byproduct is that internal constraint forces
(non-contributing forces) imposed by objects such as pin or sliding joints no
longer need be considered. The reason for this is that partial velocities
are \textit{by construction} perpendicular to forces of constraint. Constraint
forces can be included in the formulation, but will not appear in the final
equations of motion because of this very desirable property.

In summary, Kane's method allows velocities to be defined in terms of
generalized speeds, leading to simpler velocity expressions and equations of
motion expressed in first order form.  The dot product of each partial velocity
(the direction of motion associated with each generalized speed) with Newton's
second law (or Euler's equations) removes all terms which are not related to
the rate of change of each generalized speed.  This generally results in simple
equations of motion, which are easier to form.

@@

\section{Dynamics Example}
\label{sec:example}
To illustrate Kane's method in the presence of configuration and velocity
constraints, we consider a thin disk $C$ rolling without slip on a horizontal
plane. The dextral unit vectors $\hat{c}_x, \hat{c}_y, \hat{c}_z$ are fixed to $C$
with $\hat{c}_y$ perpendicular to the disk plane. Let $r$ and $m$ be the
disk radius and mass, and assume a thin disk mass distribution so that
$I_{xx}=I_{zz}=mr^2/4$ and $I_{yy} = mr^2/2$. Take the inertial (Newtonian)
frame $N$ to be similarly equipped with dextral unit vectors $\hat{n}_x,
\hat{n}_y, \hat{n}_z$, with $\hat{n}_z$ perpendicular to the ground plane
(downwards) and aligned with the local gravitational field. To orient the disk,
first align $C$ with $N$, then apply successive body fixed ZXY rotations:
yaw $q_1$, roll $q_2$, and spin $q_3$.  Denote by $A$ and $B$ the two frames
which persist after the first two intermediate rotations in the sequence; we
refer to these as the instantaneous yaw and lean frames, respectively. The unit
vector, $\hat{b}_z$ fixed in the $B$ frame, is parallel to the disk plane and
aligned with the position vector from the disk center $C^*$ to the lowest point
of the disk $P^*$. Figure \ref{fig:rolling_disk} shows how the generalized
coordinates are used to configure the disk.

@@

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.8\textwidth]{disk_drawing_cropped.pdf}
    \caption{Configuration of rolling disk using non-minimal coordinates.}
    \label{fig:rolling_disk}
\end{figure}

@@

At this point, the reader familiar with the rolling disk problem may be
wondering why we have purposefully chosen a non-minimal set of coordinates and
speeds.  In this simple example, there is an {\bf \textit{obvious and minimal}}
choice of coordinates and speeds ($q_1, q_2, q_3, q_4, q_5, u_1, u_2, u_3$
should be independent as this choice will be valid for all parameters and
configurations.  The equations of motion can be formed without ever introducing
$q_6, u_4, u_5, u_6$).

For some systems there is not an obvious and minimal choice of coordinates and
speeds (e.g., Stewart's platform, bicycle).  The values of system parameters
and the configurations in which the system will operate determine which
coordinates and speeds can be taken to be dependent (and hence potentially
eliminated by algebraic manipulations).  However, it may not be possible
\textit{a priori} to determine easily which coordinates and speeds should be
taken as dependent. Further, some systems may operate in distinct enough
regions of the configuration space that one \textit{must} use different choices
of dependent coordinates and/or speeds as the system moves from one region to
another. The purpose of our complex choice of coordinates and speeds is to
illustrate, in the context of a familiar example, how constraints must be taken
into account when linearizing equations of motion. It is our hope that by doing
this for a well known and relatively simple example, readers can apply the same
techniques to more complicated systems where its use is more appropriate or
necessary.

@@

The kinematic differential equations relate the time derivatives of the
coordinates and the generalized speeds. They are obtained by equating the
velocities written using time-differentiated generalized coordinates and the
velocities written using generalized speeds. From the addition theorem for
angular velocity, the angular velocity of $C$ is
\begin{equation}
  \label{w_qdot}
  {^N}\bar{\omega}^C = \dot{q}_1 \hat{n}_z + \dot{q}_2 \hat{a}_x + \dot{q}_3 \hat{b}_y
\end{equation}
while time differentiating $\bar{r}^{C^*}$ in $N$ yields
\begin{equation}
  \label{v_qdot}
  {^N}\bar{v}^{C^*} = \dot{q}_4 \hat{n}_x + \dot{q}_5 \hat{n}_y + \dot{q}_6 \hat{n}_z
\end{equation}
Equating (\ref{w_u}) with (\ref{w_qdot}) and (\ref{v_u}) with (\ref{v_qdot})
and expressing these vector equations in the $B$ frame, and rearranging so
that all terms appear to the left of the equality sign, we obtain
\begin{align}
    \label{rd:f_0_f_1}
\underbrace{\left[\begin{matrix}\dot{q}_{2}\\s_2
    \dot{q}_{1} + \dot{q}_{3}\\ c_2
    \dot{q}_{1}\\\dot{q}_{4}\\\dot{q}_{5}\\\dot{q}_{6}\end{matrix}\right]}_{\mathbf{f}_0}
    +
\underbrace{\left[\begin{matrix}- u_{1}\\- u_{2}\\- u_{3}\\
\left(s_1 s_2 s_3 - c_1c_3\right)u_4 +
s_1 c_2 u_5 - \left(s_1 s_2c_3 +
s_3c_1\right)u_6 \\
%
\left(-s_1c_3 - s_2 s_3c_1\right)u_4 -
c_1c_2 u_5 +
\left(-s_1 s_3 + s_2c_1c_3\right)u_6
\\
%
s_3c_2u_4  - s_2u_5 - c_2c_3u_6
\end{matrix}\right]}_{\mathbf{f}_1}
    = \left[\begin{matrix} 0\\ 0\\ 0\\ 0\\ 0\\ 0\end{matrix}\right]
\end{align}
The translational acceleration of $C^*$ and the angular acceleration of $C$,
relative to $N$, are
\begin{align}
    {^N}\bar{a}^{C^*} =& (-(u_1 s_3 + u_3 c_3) u_5 + u_2 u_6 +
                         \dot{u}_4) \hat{c}_x \notag \\
                      & + ((u_1 s_3 + u_3 c_3) u_4 - (u_1
                         c_3 - u_3 s_3) u_6 + \dot{u}_5) \hat{c}_y
                         \notag\\
                      & + ((u_1 c_3 - u_3 s_3) u_5 - u_2 u_4 +
                         \dot{u}_6) \hat{c}_z \\
    {^N}\bar{\alpha}^{C} =& (-u_2 u_3 + u_3^2 t_2 + \dot{u}_1) \hat{b}_x
                         + \dot{u}_2 \hat{b}_y + (u_1 u_2 - u_1 u_3 t_2 +
                         \dot{u}_3) \hat{b}_z
\end{align}

@@

The generalized active forces and torques for this example are
\begin{align}
    \bar{R}_{C^*} &= m g \hat{n}_z \\
    \bar{T}_C &= 0
\end{align}

@@

The twelve unknown quantities appearing in the nine kinematic and dynamic
differential equations are $\dot{q}_{1-6}$ and $\dot{u}_{1-6}$. The three
acceleration constraint equations provide the final three equations which allow
all twelve unknowns to be determined. Combining (\ref{rd:f_0_f_1}),
(\ref{rd:f_2_f_3}), and (\ref{rd:f_a}), the complete set of motion equations
can be written.
\begin{align}
\label{rd:ode}
\begin{bmatrix} \mathbf{f}_0 + \mathbf{f}_1\\
                \mathbf{f}_2 + \mathbf{f}_3\\
                \mathbf{f}_a \end{bmatrix} = \mathbf{0}
\end{align}

The naive approach to linearizing these equations of motion would be to solve
for the ($\dot{q}_i$, $\dot{u}_i$) terms to construct the right-hand side of
\begin{align}
\begin{bmatrix}\dot{\mathbf{q}} \\ \dot{\mathbf{u}}\end{bmatrix} =
    \mathbf{f}(\mathbf{q}, \mathbf{u}, t)
\end{align}
and calculate the Jacobian of $\mathbf{f}$ with respect to $[\mathbf{q} \quad
\mathbf{u}]^T$. However, this will give incorrect results.

As discussed in \cite{Schwab2003}, when $m, r,$ and $g$ are taken to be unity,
and the system is linearized about the upright steady rolling condition, the
critical speed is $v \triangleq -r\dot{q}_3 =\pm\frac{1}{\sqrt{3}}$. When
following the naive approach, and evaluating the Jacobian at the same operating
conditions and parameters, eight of the twelve eigenvalues are identically zero
while the remaining four are:
\begin{align}
  \label{eq:evals_incorrect}
  \lambda_{1,2}=\pm\frac{\sqrt{6}}{3}\sqrt{-v^2},
%  \quad \lambda_{3,4} = \pm\sqrt{\frac{4}{5} - \frac{12}{5} v^2}
  \quad \lambda_{3,4} &= \pm \frac{2\sqrt{1 - 3v^2}}{\sqrt{5}}
\end{align}

@@

Further, the fact that twelve eigenvalues are obtained should be an alert that
something is procedurally incorrect, since the number of independent quantities
is only eight (five coordinates and three speeds). We now outline a
linearization procedure which addresses this issue; we revisit this example in
Section
\ref{sec:example_revisited}.

\section{Derivation of linearization procedure}
\label{sec:derivations}
In response to the need we have demonstrated, we present a linearization
procedure that properly accounts for system constraints. Computing the Jacobian
of the right hand side of the ODEs as we did at the end of the previous
section is incorrect because it fails to apply the chain rule and thereby
properly account for the relationships imposed by configuration, velocity, and
acceleration constraints. That it isn't immediately obvious that the chain rule
needs to be applied is a byproduct of the commonly used notation which doesn't
make it explicitly clear that dependent coordinates and dependent speeds are not
only functions of time, but also functions of the independent coordinates and
independent speeds. These dependent terms should in fact be written as
\begin{align}
\label{eq:q_d_redefined}
\mathbf{q}_d (t) \to \mathbf{q}_d (\mathbf{q}_i, t) \\
\label{eq:u_d_redefined}
\mathbf{u}_d (t) \to \mathbf{u}_d (\mathbf{q}_i, \mathbf{u}_i, t)
\end{align}
While this might seem obvious, no previous author appears to make this
explicitly clear when presenting techniques for symbolically linearizing
equations of motion which are subject to constraints. While the concept is
simple in principle, correctly accounting for all quantities is tedious and
error prone. Having a high level, systematic procedure that can be implemented
reliably in software is therefore a strong argument for the detailed and
explicit procedure we present.
